AI-Based Tool for Curriculum-Based Course Timetabling of University of Potsdam

### 2024-02-08 paper.tex
polishing

\begin{abstract}
  \begin{itemize}%
    \item%
    Education is a fundamental part of universities and comes together with the time consuming non-trivial task of finding suited timetables for teaching. 
    \item%
    The corresponding problem is called Curriculum-Based Course Timetabling and considers possible constraints when looking for a conflict free timetable.
    \item% 
    We addressed this problem by following a knowledge representation and reasoning approach, describing constraints and possible assignments of courses to time slots. 
    \item% 
    Possible assignments are collected by a prototype webinterface and corresponding constraints are respected by finding automatically a timetable using an AI. 
    \item%  
    We discuss resolved redundancies, saved time and possible future benefits, when comparing traditional and new approach regarding needs of University of Potsdam. 
  \end{itemize}%
\end{abstract}

@@

\begin{itemize}%
  \item% 
  Education is a fundamental part of universities, usually addressed by offering a variety of courses may containing lectures, exercises, seminars, projects, etc. within a defined study program. 
  \item% 
  In order to obtain optimal timetables, all offered courses must be assigned to a time slot and room, respecting some constraints. 
  \item%
  A time slot is a combination of a day and a time.  
  \item%
  The corresponding problem is called Curriculum-Based Course Timetabling (\CBCTT{}; \cite{becaroto15}) and considers possible constraints when looking for a conflict free timetable. 
  \item%
  A solution to a \CBCTT{} problem is an assignment of courses to time slots and rooms, such that each corresponding hard constraint is satisfied. 
  \item% 
  An optimal solution is a solution with minimal penalty regarding soft constraints. 
  \item% 
  Traditionally, such \CBCTT{} problems are solved by humans and involving a lot of time and agreements. 
  \item% 
  Especially, when several people are involved to develop timetables, redundancies and conflicts among resulting plans may caused. 
  \item% 
  In addition, obtained timetables are often not proved to be optimal and possibly violate soft or hard constraints. 
  \item% 
  In our work, we model corresponding soft and hard constraints in context of University of Potsdam (\UP{}), to get conflict free and optimal timetables derived by an AI. 
  \item%
  Thanks to its simple modeling language and high performance solving capabilities, Answer Set Programming (\ASP{}; \cite{baral02a}) is a well suited approach of knowledge representation and reasoning to model \CBCTT{} problems, as \cite{bainkaokscsotawa18a} showed.
  \item% 
  We developed a prototype webinterface allowing planners and lecturers to enter availabilities of teaching times and suited rooms regarding courses, to find during a couple of seconds automatically a timetable provably respecting needed constraints among curricula. 
  \item% 
  By comparing the new and traditional approach, we illustrate resolved redundancies, saved time and possible future benefits. 
\end{itemize}%

@@

\begin{itemize}% 
  \item% 

@@

  \item% 
  To this end, let us first describe the structure and relation among curricula, modules, examination tasks and courses. 
  \item%

@@

  \item% 

@@

  \item% 
  Modules define the number and type of courses and course components as well as their secondary and primary examination tasks. 
  \item%
  For simplicity, in the following we refer courses and course components by courses and make explicit when needed. 
  \item% 

@@

  \item%  
  Since modules are associated to possibly several curricula, the linkage of examination tasks to courses induces involved curricula and cohorts possibly participating to a course. 
  \item% 
  Whenever a module occurs in several curricula, possibly of different institutes or faculties, then the goal to find a conflict free timetable becomes more complicated. 
  \item% 

@@

  \item% 
  The resulting timetable regarding big lectures, needs a lot of communication and agreements among several planners to ensure a conflict free timetable for involved cohorts and lecturers. 
  \item% 
  Especially, in case of degree programmes to become a teacher of allowed combinations of topics, \UP{} developed a time frame model to prevent conflicts among courses of most common combination of topics\footnote{https://www.uni-potsdam.de/en/studium/studying/organizing-your-studies/potsdams-time-frame-model}.
  \item% 
  On top of scheduled big lectures, remaining courses are planned by planners of each institute to achieve suitable timetables. 
  \item% 
  A timetable should avoid conflicts regarding multiple usage of rooms, lecturers and cohorts as well as respecting particular demands of courses, like technical equipment.
  \item% 
  A timetable becomes more optimal, if it respects for instance traveling time of consecutive courses, teaching load per day and gaps among courses of a day aiming on same cohort or lecturer, respectively. 
  \item% 
  As soon as a solution is found, each planner has to copy it course by course to universities course catalog, link each course to examination tasks of particular modules and book a room for the assigned time slot. \\
  \item% 
  Our approach aims on time consuming communication among planners and lecturers as well as redundancies when a solution was found but has to get copied to the course catalog, linked to modules and booked to rooms. 
  \item% 
  To this end, we developed a prototype webinterface allowing planners and lecturers to create courses and course components together with its module linkages, prioritized available time slots as well as rooms and relations among courses. 
  \item% 
  Four pairwise relations among courses are provided:
    \item two course components are not allowed to take place in parallel, e.g. obligatory lectures of same cohort
    \item two course components have to take place in parallel (same time slot, but different rooms), e.g. two groups of exercises of the same course
    \item two course components have to take place simultaneous (same time slot and room), e.g. hosting two seminars sharing major content in same room
    \item two course components have to take place consecutive, e.g. having an exercise directly after the corresponding lecture 
  \item%
  For each course component, the webinterface allows to state the expected number of participants as well as the need of particular equipment, e.g. large board.  
  \item% 
  See figure~\ref{fig:webinterface} illustrating parts of the prototype webinterface. 

@@

    \caption{Webinterface inducing availabilities by the cross product of days, times and rooms for a lecture.}
    \label{fig:webinterface}

@@

  \item%
  The corresponding planner checks the consistence and correctness of the input data for the webinterface. 
  \item% 
  The input data together with a logic representation of necessary constraints is given to an AI, that searches for an optimal timetable. 
  \item% 
  The solution is a visualization of the resulting timetable and double checked by planners. 
  \item% 
  Note that the new approach follows the traditional idea of splitting the process into two steps, by first finding a timetable regarding big lectures and then completing it to a timetable for each institute. 
  \item% 
  The resulting timetable of each institute has to (automatically) transferred to universities course catalog respecting linked modules, time slots and rooms. 
  \item% 
  This last step of integrating the obtained timetable automatically to the course catalog is not yet part of the prototype but will omit a lot of redundancies. 
\end{itemize}%

%------------------------------------------------------------------------------
\section{\ASP{} Paradigm}
\label{sec:asp}
\begin{itemize}%
  \item% 
  \ASP{} \cite{baral02a, breitr11a} is an approach to declarative problem solving and belongs to the area of knowledge representation and reasoning. 
  \item% 
  The roots of \ASP{} go back to logic programming, nonmonotonic reasoning and constraint satisfaction \cite{gellif91a}. 
  \item% 
  The idea of \ASP{} is to describe the problem by using a formal representation, rather than telling a computer how to solve the problem. 
  \item% 
  Figure~\ref{fig:asp} illustrates this approach \cite{gekakasc12a}. 
  \begin{figure}
  \label{fig:asp}
  \begin{center}{%
    \begin{picture}(300,120)(-150,-60)
    \put(-80,+40){\makebox(0,0){\framebox(80,20){Problem}}}
    \put(-80,-40){\makebox(0,0){\framebox(80,20){Logic Program}}}
    \put(+80,+40){\makebox(0,0){\framebox(80,20){Solution}}}
    \put(+80,-40){\makebox(0,0){\framebox(80,20){Stable Models}}}
    \put(-80,+30){\vector(0,-1){60}}
    \put(-40,-40){\vector(+1,0){80}}
    \put(+80,-30){\vector(0,+1){60}}
    \put(-110,  0){\makebox(0,0){Modeling}}
    \put(+120,  0){\makebox(0,0){Interpreting}}
    \put(   0,-55){\makebox(0,0){{Solving}}}
    \end{picture}}
    \end{center}
  \caption{Declarative problem solving.}
  \end{figure}
  \item% 
  An easy and human readable modeling language \cite{cafageiakakrlemarisc19a} is used to obtain a logic program modeling the original problem. 
  \item%
  By using problem solvers like \clingo{} \cite{gekakaosscwa16a}, a solution, i.e. stable model \cite{gellif88b}, of a problem is found and has to be interpreted by the user. 
  \item% 
  Due to its simple but rich high-level modeling language as well as its elaboration-tolerance and its high-performance solving capacities, \ASP{} becomes a well suited approach to model real world problems. 
  \item% 
  The range of applications to model and solve combinatorial search problems is among planning, scheduling, configuration, probabilistic reasoning, diagnosis and repair, classification, query answering, explanation generation, multi-agent systems, natural language processing, computational biology, music composition, model checking and robotics \cite{ergele16}. 
  \item% 
  In Terms of industrial applications, \ASP{} has been used for solving (automated product) configuration \cite{soinie99a, hebasasc22a}, planning \cite{jizhkhst19a, sopobasc23a}, scheduling \cite{aldoma17a}, timetabling \cite{bainkaokscsotawa18a} and other problems \cite{fafrsctate18a}. 
  \item%
  Especially, for industrial applications, tackled by a declarative approach like \ASP{}, significant improvements wrt implementation, maintenance costs and human-machine interaction were shown. 
  \item% 
  For instance, Siemens reported to save more than 80\% of maintenance costs by using a knowledge representation and reasoning approach \cite{fafrhascsc16a}. 
\end{itemize}%

@@

\begin{itemize}%
  \item%
  \ASP{} provides a powerful high-level modeling language, which allows recursive definitions, default negation for dealing with the absence of information, disjunctions, aggregates, weight constraints, optimization statements and external atoms. 
  \item%
  In the following, we give a short introduction to basics of the modeling language of \ASP{} \cite{cafageiakakrlemarisc19a} and the stable models semantics of logic programs \cite{gellif88b}. \\
  \item% 

@@

  \item% 
  Default negation, also known as negation as failure, refers to the absence of information, whereas "classical" negation induces the presence of negated information.  
  \item% 

@@

  \item% 
  A rule $r$ is called (integrity) constraint, whenever $\head{(r)}\cap\body{(r)}^-\neq\emptyset$.%\footnote{Usually, literals $\head{(r)}$ and $\naf{} \head{(r)}$ are dropped in this case.}
  \item% 
  Integrity constraints are used to eliminate unintended solution candidates, whenever its body is satisfied. 
  \item% 

@@

  \lstinputlisting[basicstyle=\small, firstline=27, lastline=31]{example.lp}
  deriving a conflict among two course components of ``AI'' lecture and exercise that shall take place consecutive, whenever they are scheduled on different days.\footnote{In the syntax of an \ASP{} encoding `$\leftarrow$' is represented by `$:-$' and each rule terminates by a period `.'.}  
  \item%

@@

    P^X=\{\head{(r)}\leftarrow \body{(r)}^+\mid r\in P, \body{(r)}^-\cap X = \emptyset\}

@@

  \item% 
  $X$ is a stable model of a program $P$, if $X$ is the $\subseteq$-minimal model of $P^X$. 
  \item% 
  As an example consider the following program $P$
  \lstinputlisting[basicstyle=\small, numbers=left]{example.lp}

@@

  \item% 
  Since atoms of lines 1-21 are facts, they belong to any model of $P$. 
  \item% 

@@

  \item% 
  Note that $X$ is not a model of $P$, since the body of rule in line 22 is satisfied, but not its head. 
  \item% 

@@

  are models of $P$, but the latter one is the only stable model of $P$. \\
  \item% 

@@

  \item%  
  In case of \CBCTT{}, one could think of generating any timetable as a model and ruling out certain models by posing particular (hard) constraints. 
\end{itemize}%

\subsection{Modeling \CBCTT{} of \UP{}}
\label{subsec:modeling}
\begin{itemize}%
  \item% 
  Since a course may consists of several course components of possibly different types, e.g. a lecture and an exercise, we identify a course component by a unique tuple of a name and its type, e.g. lecture \textit{(``Artificial Intelligence'', v)} and exercise \textit{(``Artificial Intelligence'', u)} of Artificial Intelligence course. 
  \item%
  We modeled the following hard (H0-8) and soft (S0-13) constraints.

@@

  \item% 
  Note that some constraints are similar to those presented in \cite{bainkaokscsotawa18a}, others are new or modified.
  \item% 
  Due to lack of space, we drop the corresponding logic program here\footnote{The particular \ASP{} encoding representing above constraints can be found at GitHub: https://github.com/schellhorn/CBCTT-UP.}.  
  \item% 
  Optimization tries to minimize penalty points regarding above soft constraints. 
  \item%
  The system of \cite{bainkaokscsotawa18a}, reads instances of a standard input format \cite{bocegasc12a}, translates them into \ASP{} facts and assigns potentially any course to any time slot. 
  \item%
  In contrast, we used a direct modeling dedicated to constraints and needs of the \UP{}, which partially is not covered by the standard input format. 
  \item%
  As a design decision, we reduced the search space to collected availabilities only, instead of checking for all possible time slots and rooms for each course component.
\end{itemize}%

%------------------------------------------------------------------------------
\section{Results and Impact}
\label{sec:impact}
\begin{itemize}%
  \item%
  In \cite{bainkaokscsotawa18a} was shown, that treating \CBCTT{} problems by an \ASP{} based approach outperforms state of the art approaches and makes it well suited to tackle \CBCTT{} problems. 
  \item% 
  Our real world approach is still at a prototype stage and is lacking access to some data needed to apply all presented and modeled constraints. 
  \item%
  For winter 2023/24, an optimal and conflict free plan was found for about 160 courses aimed at big lecture halls of the Faculty of Science, all 90 courses of the Institute of Mathematics and 100 courses of the Institute of Computer Science, involving constraints H1, H3, H5, S0, S1 and S7 of above \ASP{} based modeling. 
  \item% 
  Analogously, for summer 2024, an optimal and conflict free plan was found involving about 200 courses and constraints H0, H1, H3, H5, H6, H7, H8, S0, S1, S4 and S7.  
  \item% 
  Solver \clingo{} was integrated to the web interface and needed about 170 seconds for winter and less than one second for summer, respectively, to solve the problem regarding the faculty. 
  \item% 
  Finding plans for the institutes took less than a second, respectively. 
  \item% 
  A major benefit of the presented AI-based approach is its fairness, transparency and elaboration tolerance. 
  \item% 
  Planners and lecturers are able to set a prioritized selection of availabilities and preferences to upcoming teaching periods, resulting in a provable conflict free timetable. 
  \item% 
  Getting a conflict free timetable for a faculty in a couple of seconds, respecting dependencies among several institutes at the same time, resolves a lot of communication overhead and conflicts of the earlier approach. 
  \item% 
  Thus, our approach gives planners already the opportunity to address their additional available time to other tasks. 
  \item% 
  As soon as the integration of the resulting timetable to the universities course catalog is done automatically, then even more redundancies are resolved and a lot more capacities become available. 
  \item% 
  Due to our current prototype state, we did not yet compare to other approaches\cite{feutrier23a} regarding performance. 
\end{itemize}%

@@

%------------------------------------------------------------------------------
\section{Future Work}
\label{sec:future}
\begin{itemize}%
  \item% 
  For the future, we plan to involve an interface, allowing us to access data like previous course catalogues and curricula as well as their modules.
  \item% 
  Furthermore, we plan to develop an interface, allowing us to push final timetables to universities course catalog automatically. 
  \item% 
  Another, not trivial task, is to provide opportunities to the user to analyze and solve conflicts, whenever needed.  
  \item% 
  In a next step, we are going to extend the number of users and involve them in a feedback loop, to improve the usage of the system and check for additional needs. 
  \item% 
  In a long run, one could think about to add functionalities like drag and drop to take smaller changes by hand to a calculated timetable into account, preserving consistence regarding hard and soft constraints. 
  \item% 
  When establishing the mentioned workflow to get conflict free timetables, one could measure the average amount of time needed per planner using the traditional and new approach, respectively.
  \item% 
  Moreover, one could compare the quality of resulting timetables regarding desired hard and soft constraints. 
  \item% 
  Comparing quality and used time of both approaches may highlight the usage and impact of our new approach using AI of knowledge representation and reasoning. 
  \item% 
  The data about availabilities of different types of course components to several time slots, provides insights to the needs of room infrastructure. 
  \item% 
  One could think about, to possibly exploit entered availabilities and number of participants to derive location, size and quantity of needed lecture halls and seminar rooms, when plan to build new ones.
  \item% 
  Finally, our approach provides a lot potential to get rid of redundancies, save money and lift administrative planning tasks to the next level. 
\end{itemize}%

\label{sec:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{refs}

>>>>

\begin{abstract}
  %\begin{itemize}%
    %\item%
    Education is a fundamental component of universities and is intricately linked with the time-consuming non-trivial task of creating suitable timetables for teaching. 
    %\item%
    The corresponding problem is known as Curriculum-Based Course Timetabling, which takes into account various constraints while seeking a conflict-free timetable. \\
    %\item% 
    We addressed this problem by following a knowledge representation and reasoning approach, describing constraints and feasible assignments of courses to time slots. 
    %\item% 
    Possible assignments are gathered through a prototypical web interface, and the corresponding constraints are respected by finding automatically a timetable using an AI. \\
    %\item%  
    We discuss the resolved redundancies, time saved and potential future benefits when comparing the traditional and new approach regarding the needs at the University of Potsdam. 
  %\end{itemize}%
\end{abstract}

@@

%\begin{itemize}%
  %\item% 
  Education is a fundamental component of universities, usually addressed by offering a variety of courses that may include lectures, exercises, seminars, projects, etc., within a defined study program. 
  %\item% 
  In order to obtain optimal timetables, all offered courses must be assigned to a time slot and room, while respecting some hard and soft constraints. 
  %\item%
  A time slot is a combination of a day and a time. \\
  %\item%
  The associated problem is known as Curriculum-Based Course Timetabling (\CBCTT{}; \cite{becaroto15}), which takes into account various constraints when seeking a conflict-free timetable. 
  %\item%
  A solution to a \CBCTT{} problem is an assignment of courses to time slots and rooms, such that each corresponding hard constraint is satisfied. 
  %\item% 
  An optimal solution is a solution with minimal penalties regarding soft constraints. \\
  %\item% 
  Traditionally, solving \CBCTT{} problems involve human efforts, consuming a significant amount of time and requiring numerous agreements.
  %\item% 
  Especially when multiple individuals are involved in developing timetables, redundancies and conflicts among the resulting plans may arise.
  %\item% 
  In addition, the obtained timetables are often not guaranteed to be optimal and may possibly violate soft or hard constraints. \\
  %\item% 
  In our work, we model the relevant soft and hard constraints within the context of the University of Potsdam (\UP{}) to obtain conflict-free and optimal timetables derived by an AI. 
  %\item%
  Thanks to its simple modeling language and high-performance solving capabilities, Answer Set Programming (\ASP{}; \cite{baral02a}) is a well-suited approach for knowledge representation and reasoning in modeling \CBCTT{} problems, as demonstrated by \cite{bainkaokscsotawa18a}.
  %\item% 
  We developed a prototypical web interface allowing planners and lecturers to enter availabilities for teaching times and suited rooms regarding courses. 
  %\item% 
  This interface can automatically generate a timetable within seconds, ensuring compliance with the required constraints among curricula. 
  %\item%
  The idea is to find a solution from the time slots specified by lecturers rather than permitting any course to potentially take place at any time slot.
  %\item%
  As a goal, the obtained timetables should automatically result into according room reservations at universities course catalog.  
  %\item% 
  In comparing the new and traditional approach, we sketch resolved redundancies, time savings and possible future benefits. 
%\end{itemize}%

@@

%\begin{itemize}% 
  %\item% 

@@

  %\item% 
  To this end, let us first describe the structure and relationships among curricula, modules, examination tasks and courses. \\
  %\item%

@@

  %\item% 

@@

  %\item% 
  Modules specify the number and type of courses and course components, along with their secondary and primary examination tasks. 
  %\item%
  For simplicity, in the following, we use the term \textit{courses} to refer to both courses and course components, explicitly indicating the distinction when necessary. 
  %\item% 

@@

  %\item%  
  As modules are associated with possibly several curricula, the linkage of examination tasks to courses induces involved curricula and cohorts possibly participate to a course. 
  %\item% 
  Whenever a module occurs in several curricula, possibly of different institutes or faculties, then the objective to find a conflict-free timetable becomes more complicated. \\
  %\item% 

@@

  %\item% 
  The resulting timetable regarding big lectures, needs a lot of communication and agreements among several planners to ensure a conflict-free timetable for involved cohorts and lecturers. 
  %\item% 
  Especially, in the case of degree programmes that aim to train teachers with allowed combinations of topics, \UP{} has developed a time frame model to prevent conflicts among courses of most common combinations of topics\footnote{https://www.uni-potsdam.de/en/studium/studying/organizing-your-studies/potsdams-time-frame-model}.
  %\item% 
  On top of scheduled big lectures, the remaining courses are planned by the respective planners of each institute to achieve suitable timetables. 
  %\item% 
  A timetable should avoid conflicts regarding multiple use of rooms, lecturers and cohorts as well as respecting particular demands of courses, like technical equipment.
  %\item% 
  A timetable is considered more optimal, if it respects for instance traveling time of consecutive courses, teaching load per day and gaps among courses of a day aiming on same cohort or lecturer, respectively. 
  %\item% 
  Once a solution is found, each planner must manually to copy it, course by course, to the universities course catalog, link each course to examination tasks of particular modules and book a room for the assigned time slot. \\
  %\item% 
  Our approach aims to minimize time-consuming communication among planners and lecturers, as well as reduce redundancies when a solution is found but needs to be copied to the course catalog, linked to modules, and booked into rooms.
  %\item% 
  To this end, we developed a prototypical web interface that enables planners and lecturers to collaboratively create courses and course components, providing all necessary information to facilitate finding a feasible solution.
  %\item%
  In more detail, for each course, the system requests a name, a list of linked modules, specification of whether it is offered in the winter or summer term, and the designation of a responsible planner.
  %\item% 
  The web interface allows to state the type, e.g. lecture or seminar, the expected number of participants and any specific equipment requirements, e.g. large board, for each course component.  
  %\item%
  Furthermore, for each course component, particular rooms, days and times, along with a priority can be selected.
  %\item% 
  Through  the cross product of days, times and rooms possible prioritized availabilities are established for each course component. 
  %\item% 
  See figure~\ref{fig:web interface} illustrating parts of the prototypical web interface. 

@@

    \caption{Web interface inducing 18 availabilities of highest priority 1 and three availabilities of priority 2 for a seminar with expected number of 30 participants.}
    \label{fig:web interface}

@@

  %\item% 
  We identified and provided four pairwise relations to capture significant constraints among course components:
    \item Two course components are not allowed to take place in parallel, e.g. obligatory lectures for the same cohort.
    \item Two course components have to take place in parallel (same time slot, but different rooms), e.g. two groups of exercises for the same course.
    \item Two course components have to take place simultaneous (same time slot and room), e.g. hosting two seminars sharing major content in the same room.
    \item Two course components have to take place consecutive, e.g. having an exercise directly after the corresponding lecture. 
  %\item%
  These four relations allow us to pose certain structures among course components, which enable or disable underlying constraints as elaborated in the next section. \\ 
  %\item%
  The corresponding planner verifies the consistency and correctness of the input data for the web interface. 
  %\item% 
  The input data, along with a logic representation of necessary constraints, is given to an AI problem solver, which then searches for an optimal timetable. 
  %\item% 
  The output is a visualization of the resulting timetable and is double-checked by planners. 
  %\item%
  As an example of a resulting timetable see figure~\ref{fig:timetable}. 
    \includegraphics[width=\linewidth]{timetable}
    \caption{Resulting timetable containing about 200 course components of the Faculty of Science for summer 2024.}
    \label{fig:timetable}
  \end{figure} \\
  %\item% 
  Note that the new approach aligns with the traditional idea of splitting the process into two steps: initially finding a timetable for large lectures and subsequently completing it to a timetable for each institute.
  %\item% 
  The resulting timetable for each institute must be (automatically) transferred to the university's course catalog, taking into account linked modules, lecturers, time slots, and rooms.
  %\item% 
  The final step of automatically integrating the obtained timetable into the course catalog is not yet implemented in the prototype but will eliminate many redundancies.
%\end{itemize}%

%------------------------------------------------------------------------------
\section{\ASP{} Paradigm}
\label{sec:asp}
%\begin{itemize}%
  %\item% 
  \ASP{} \cite{baral02a, breitr11a} is an approach to declarative problem solving and belongs to the area of knowledge representation and reasoning. 
  %\item% 
  The roots of \ASP{} go back to logic programming, nonmonotonic reasoning and constraint satisfaction \cite{gellif91a}. 
  %\item% 
  The idea of \ASP{} is to describe the problem using a formal representation, rather than instructing a computer how to solve the problem. 
  %\item% 
  Figure~\ref{fig:asp} illustrates this approach \cite{gekakasc12a}. 
  \begin{figure}
  \label{fig:asp}
  \begin{center}{%
    \begin{picture}(300,120)(-150,-60)
    \put(-80,+40){\makebox(0,0){\framebox(80,20){Problem}}}
    \put(-80,-40){\makebox(0,0){\framebox(80,20){Logic Program}}}
    \put(+80,+40){\makebox(0,0){\framebox(80,20){Solution}}}
    \put(+80,-40){\makebox(0,0){\framebox(80,20){Stable Models}}}
    \put(-80,+30){\vector(0,-1){60}}
    \put(-40,-40){\vector(+1,0){80}}
    \put(+80,-30){\vector(0,+1){60}}
    \put(-110,  0){\makebox(0,0){Modeling}}
    \put(+120,  0){\makebox(0,0){Interpreting}}
    \put(   0,-55){\makebox(0,0){{Solving}}}
    \end{picture}}
    \end{center}
  \caption{Declarative problem solving.}
  \end{figure} \\
  %\item% 
  An easy and human-readable modeling language \cite{cafageiakakrlemarisc19a} is used to create a logic program modeling the original problem. 
  %\item%
  By utilizing a problem solver such as \clingo{} \cite{gekakaosscwa16a}, a solution, i.e. a stable model \cite{gellif88b}, of the problem is found and has to be interpreted by the user. 
  %\item% 
  Due to its simple but rich high-level modeling language, its elaboration-tolerance, and its high-performance solving capacities, \ASP{} becomes a well suited approach to model real-world problems. \\
  %\item% 
  The range of applications for modeling and solving combinatorial search problems with \ASP{} spans across various domains, including planning, scheduling, configuration, probabilistic reasoning, diagnosis and repair, classification, query answering, explanation generation, multi-agent systems, natural language processing, computational biology, music composition, model checking and robotics \cite{ergele16}. 
  %\item% 
  In Terms of industrial applications, \ASP{} has been successfully employed for solving (automated product) configuration \cite{soinie99a, hebasasc22a}, planning \cite{jizhkhst19a, sopobasc23a}, scheduling \cite{aldoma17a}, timetabling \cite{bainkaokscsotawa18a} and other problems \cite{fafrsctate18a}. 
  %\item%
  Especially for industrial applications, tackled by a declarative approach like \ASP{}, significant improvements wrt implementation, maintenance costs and human-machine interaction were shown. 
  %\item% 
  For instance, Siemens reported saving more than 80\% of maintenance costs by using a knowledge representation and reasoning approach, as mentioned in \cite{fafrhascsc16a}. 
%\end{itemize}%

@@

%\begin{itemize}%
  %\item%
  \ASP{} provides a powerful high-level modeling language that supports recursive definitions, default negation for dealing with the absence of information, disjunctions, aggregates, weight constraints, optimization statements and external atoms. 
  %\item%
  In the following, we provide a brief introduction to the basics of the modeling language of \ASP{} \cite{cafageiakakrlemarisc19a} and the stable models semantics of logic programs \cite{gellif88b}. \\
  %\item% 

@@

  %\item% 
  Default negation, also known as negation as failure, refers to the absence of information, whereas "classical" negation induces the presence of negated information. \\
  %\item% 

@@

  %\item% 
  A rule $r$ is called (integrity) constraint, whenever $\head{(r)}\cap\body{(r)}^-\neq\emptyset$. %\footnote{Usually, literals $\head{(r)}$ and $\naf{} \head{(r)}$ are dropped in this case.}
  %\item% 
  Integrity constraints are used to eliminate unintended solution candidates, whenever its body is satisfied. \\
  %\item% 

@@

  \lstinputlisting[basicstyle=\ttfamily\small, firstline=27, lastline=31]{example.lp} %TODO: Listing style von potassco solution 
  deriving a conflict among two course components of ``AI'' lecture and exercise that shall take place consecutive, whenever they are scheduled on different days.\footnote{In the syntax of an \ASP{} encoding `$\leftarrow$' is represented by `$:-$' and each rule terminates by a period `.'.}  \\
  %\item%

@@

    P^X=\{\head{(r)}\leftarrow \body{(r)}^+\mid r\in P, \body{(r)}^-\cap X = \emptyset\}.

@@

  %\item% 
  $X$ is a stable model of a program $P$, if $X$ is the $\subseteq$-minimal model of $P^X$. \\
  %\item% 
  As an example consider the following program $P_1$
  \lstinputlisting[basicstyle=\ttfamily\small, numbers=left]{example.lp}

@@

  %\item% 
  Since atoms of lines 1-21 are facts, they belong to any model of $P_1$. 
  %\item% 

@@

  %\item% 
  Note that $X$ is not a model of $_1$, since the body of rule in line 22 is satisfied, but not its head. 
  %\item% 

@@

  are models of $P_1$, but the latter one is the only stable model of $P_1$, since first model is not $\subseteq$-minimal under reduct wrt $P_1$, due to 
         &\texttt{conflict(consecutive, ((42, ("AI", v)), (43, ("AI", u))))}\}.
  \end{align*} \\
  %\item% 

@@

  %\item%  
  In the case of \CBCTT{}, one could think of generating any timetable as a model and eliminating certain models by imposing specific (hard) constraints.
%\end{itemize}%

\subsection{Modeling \CBCTT{} of \UP{}}
\label{subsec:modeling}
%\begin{itemize}%
  %\item% 
  Since a course may consists of several course components of possibly different types, e.g. a lecture and an exercise, we identify a course component by a unique tuple of a name and its type, e.g. lecture \textit{(``AI'', v)} and exercise \textit{(``AI'', u)} of Artificial Intelligence course. \\
  %\item%
  We modeled the following hard (H0-8) and soft (S0-13) constraints.

@@

  %\item% 
  Note that some constraints are similar to those presented in \cite{bainkaokscsotawa18a}, others are modified or new.
  %\item% 
  Due to lack of space, we omit the corresponding logic program here.\footnote{The particular \ASP{} encoding representing above constraints can be found at GitHub: https://github.com/schellhorn/CBCTT-UP.} 
  %\item% 
  Optimization aims to minimize penalty points associated with the aforementioned soft constraints. \\
  %\item%
  The system of \cite{bainkaokscsotawa18a}, reads instances in a standard input format \cite{bocegasc12a}, translates them into \ASP{} facts and assigns potentially any course to any time slot. 
  %\item%
  In contrast, we employed a direct modeling approach tailored to the constraints and requirements of the \UP{}, which partially is not covered by the standard input format. 
  %\item%
  As a design decision, we reduced the search space to collected availabilities only, rather than of checking for all possible time slots and rooms for each course component.
%\end{itemize}%

%------------------------------------------------------------------------------
\section{Results and Impact}
\label{sec:impact}
%\begin{itemize}%
  %\item%
  In \cite{bainkaokscsotawa18a}, it was demonstrated that addressing \CBCTT{} problems through an \ASP{}-based approach outperforms state-of-the-art approaches and thus makes it well-suited to tackle \CBCTT{} problems. 
  %\item% 
  Our real world approach is still at a prototypical stage and is lacking access to some data needed to apply all presented and modeled constraints. \\
  %\item%
  For the winter semester of 2023/24, an optimal and conflict-free plan was found for approximately 160 courses aimed at big lecture halls of the Faculty of Science, all 90 courses of the Institute of Mathematics and 100 courses of the Institute of Computer Science, incorporating constraints H1, H3, H5, S0, S1 and S7 from the above \ASP{}-based modeling. 
  %\item% 
  Analogously, for the summer semester of 2024, an optimal and conflict-free plan was found involving about 200 courses and constraints H0, H1, H3, H5, H6, H7, H8, S0, S1, S4 and S7.  
  %\item% 
  The \clingo{} solver was integrated into the web interface and required approximately  170 seconds for the winter and less than one second for the summer semester, respectively, to solve the problem for the faculty. 
  %\item% 
  Finding plans for the institutes took less than a second, respectively. 
  %\item%
  The time required to solve these problems relative to the number of involved courses highlights the exponential complexity inherent in \CBCTT{} problems. 
  %\item% 
  For this reason, it is crucial to maintain a workflow that breaks down the planning task into several smaller tasks, as the traditional approach does. \\
  %\item% 
  A major benefit of the presented AI-based approach is its fairness, transparency and elaboration tolerance. 
  %\item% 
  Thanks to its elaboration tolerance, the existing encoding can be reused for searching a plan for each upcoming semester and depends only on given courses. 
  %\item% 
  It is possible to add new constraints or relax existing ones whenever needed to address the special needs of particular institutes or faculties. 
  %\item% 
  Planners and lecturers can set a prioritized selection of availabilities and preferences for upcoming teaching periods, resulting in an optimal and provably conflict-free timetable.
  %\item% 
  Obtaining a conflict-free timetable for a faculty in a matter of seconds, while respecting dependencies among several institutes at the same time, resolves a significant amount of communication overhead and conflicts inherent in the earlier approach. 
  %\item% 
  Thus, our approach provides planners with the opportunity to allocate their additional available time to other tasks. \\
  %\item% 
  Once the integration of the resulting timetable into the university's course catalog is done automatically, even more redundancies are resolved, and a considerable amount of additional capacities become available. 
  %\item% 
  Due to our current prototypical state, we have not yet conducted a performance comparison with other approaches\cite{feutrier23a}. 
%\end{itemize}%

@@

%------------------------------------------------------------------------------
\section{Future Work}
\label{sec:future}
%\begin{itemize}%
  %\item% 
  For the future, we plan to implement an interface that allows us to access data such as previous course catalogues, curricula, and their modules.
  %\item% 
  Furthermore, we plan to develop an interface that allows us to automatically push final timetables to the university's course catalog. \\
  %\item% 
  Another non-trivial task is to provide opportunities to the user to analyze and solve conflicts whenever needed.  
  %\item% 
  In a next step, we plan to expand the number of users and involve them in a feedback loop to enhance the usage of the system and identify additional needs.
  %\item% 
  In a long run, one could think about to add functionalities such as drag and drop to manually incorporate smaller changes to a calculated timetable, while preserving consistency regarding hard and soft constraints. \\
  %\item% 
  When establishing the mentioned workflow to obtain conflict-free timetables, one could measure the average amount of time needed per planner using the traditional and new approach, respectively.
  %\item% 
  Moreover, one could compare the quality of resulting timetables regarding the desired hard and soft constraints. 
  %\item% 
  Comparing the quality and time efficiency of both approaches may highlight the usage and impact of our new approach using AI of knowledge representation and reasoning. \\
  %\item% 
  The data about availabilities of different types of course components in various time slots provides insights into the needs of room infrastructure. 
  %\item% 
  One could consider possibly leveraging entered availabilities and the number of participants to derive the location, size, and quantity of needed lecture halls and seminar rooms when planning to build new ones. \\
  %\item% 
  Finally, our approach has significant potential to eliminate redundancies, save money, and elevate administrative planning tasks to the next level.
%\end{itemize}%

\label{sec:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{refs} % TODO: consistence eg for first names
